---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# R-package: FunctionalPrediction

<!-- badges: start -->

[![R-CMD-check](https://github.com/creutzml/FunctionalPrediction/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/creutzml/FunctionalPrediction/actions/workflows/R-CMD-check.yaml)
[![pages-build-deployment](https://github.com/creutzml/FunctionalPrediction/actions/workflows/pages/pages-build-deployment/badge.svg)](https://github.com/creutzml/FunctionalPrediction/actions/workflows/pages/pages-build-deployment)
[![pkgdown](https://github.com/creutzml/FunctionalPrediction/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/creutzml/FunctionalPrediction/actions/workflows/pkgdown.yaml)
[![test-coverage](https://github.com/creutzml/FunctionalPrediction/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/creutzml/FunctionalPrediction/actions/workflows/test-coverage.yaml)
<!-- badges: end -->


# Description 
The R package `FunctionalPrediction` allows one to compute simultaneous confidence bands
for function-valued parameters (e.g. for mean functions
$\mu=\{\mu(t):t\in[0,1]\}$) and adds the ability to create one-sided bands, 
prediction bands, and simultaneous bands for the conditional mean of a 
functional concurrent regression model. Users can install `FunctionalPrediction` directly, and it will add on the original capabilities of `ffscb` alongside the concurrent regression model bands. 

The band shapes are constructed under
the fairness constraint of *false positive rate balance* which allows to
distribute the total false positive rate $\alpha\in(0,1)$ over
partitions $0 = a_0 < a_1 < \ldots < a_p = 1$ of the function domain $[0,1]$.[^1]
This leads to band shapes that are interpretable both globally and
locally. Globally (i.e. over the total domain $[0,1]$) the simultaneous
confidence band provides a $(1-\alpha)\times 100\%$ coverage
probability. Locally (i.e. over sub-intervals
$[a_{j-1},a_j]\subseteq[0,1]$) the simultaneous confidence band provides
a $(1-\alpha(a_j-a_{j-1}))\times 100\%$ coverage probability.

* The statistical theory and methodology for simultaneous confidence bands of the functional    mean of a univariate sample are described in: <br> Fast ‘n’
  fair simultaneous confidence bands for functional parameters as
  introduced in the paper [**Fast and Fair Simultaneous Confidence Bands
  for Functional Parameters**
  (arXiv:1910.00131)](https://arxiv.org/abs/1910.00131) by [Dominik
  Liebl](www.dliebl.com) and [Matthew
  Reimherr](http://www.personal.psu.edu/mlr36/).

* The R-codes of the R package `ffscb` can be found at the GitHub repo
  <https://github.com/lidom/ffscb>

* The statistical theory and methodology for simultaneous confidence and prediction bands of    the conditional mean of a concurrent functional regression model are described in the paper   **Fair Simultaneous Prediction and Confidence
  Bands for Concurrent Functional Regressions: Comparing Sprinters with Prosthetic
  versus Biological Legs** by [Michael L.     Creutzinger](https://www.linkedin.com/in/michaellcreutzinger/),
   [Dominik Liebl](www.dliebl.com), and [Julia L. Sharp](https://sites.google.com/view/juliasharp/home).

* The R-codes of the R package `FunctionalPrediction` can be found at the GitHub repo
  <https://github.com/creutzml/FunctionalPrediction>


## Installation 
```{r, eval = FALSE, echo=TRUE}
devtools::install_github("creutzml/FunctionalPrediction")
install.packages("ggplot2")
```



## Small example based on artifical data


```{r, fig.height=5}
# Load the package and supporting packages
library(FunctionalPrediction)
library(ggplot2)

## Generate data:
#################################################################
# Number of observations and sampling points, 
n_obs <- 30
n_sp <- 101
grid <- make_grid(n_sp, rangevals = c(0, 1))
y_vals_mat <- matrix(0, nrow = n_sp, ncol = n_obs + 2)

# Need an array for the predictor matrix
x_array <- array(c(matrix(1, nrow = n_sp, ncol = n_obs + 2), 
matrix(c(rep(0, n_sp*n_obs/2),
rep(1, n_sp*n_obs/2), 
rep(0, n_sp), rep(1, n_sp)), 
nrow = n_sp, ncol = n_obs + 2)), 
dim = c(length(grid), n_obs + 2, 2))

# Beta functions
b0 <- 1
b1 <- function(i, t_mat) {
  sin(8*pi*t_mat[i])*exp(-3*t_mat[i]) + t_mat[i]
}

# True mean functionals
true_int_mean <- rep(b0, length(grid))
true_int_beta_mean <- true_int_mean + b1(1:length(grid), grid)
true_slope_mean <- b1(1:length(grid), grid)

## Generating response values
# Random error covariance with non-stationary matern covariance
cov.m <- make_cov_m(
  cov.f = ffscb::covf_nonst_matern, #covf_st_matern,
  grid = grid, 
  cov.f.params=c(2, 1/4, 1/4) #c(3/2, 1/4)
)

# t-distributed errors with nu_0 = 5
eps_mat <- make_sample(
  mean.v = rep(0, n_sp),
  cov.m = cov.m,
  N = n_obs + 2,
  dist = "rnorm"
)*sqrt(5/rchisq(1, df = 5))


# Generate the random data
y_vals_mat <- t(x_array[1,,]%*%
  t(cbind(true_int_mean, true_slope_mean))) + 
  eps_mat

# matplot(y_vals_mat, type = "l")


# Next, remove the last two columns for hold out prediction
y_vals_mat_temp <- y_vals_mat[,-c(n_obs + 1, n_obs + 2)]
x_array_temp <- x_array[,-c(n_obs + 1, n_obs + 2),]
y_vals_mat_ho <- y_vals_mat[,c(n_obs + 1, n_obs + 2)]
x_array_ho <- x_array[,c(n_obs + 1, n_obs + 2),]
#################################################################



## FFSCB fitting
#################################################################
## And now, if we fit with concurrent regression I made and our bands
fReg_list <- FunctionalPrediction::fRegress_concurrent(
  y_mat = y_vals_mat_temp, 
  x_array = x_array_temp[,,-1]
)
    
# First prediction: c(1,0)
fBands <- FunctionalPrediction::predict_concurrent(
  concurrent_list = fReg_list,
  interval = "prediction", 
  err_str = "t",
  new_dat = c(0), 
  conf.level = 0.90, 
  n_int = 3,
  nu0_hat = "singh_df",
  mse_scalar = "ub"
)

# Data frame to plot the band
band_plot_df <- data.frame(
  t = seq(0,1, length.out = 101),
  y_true = y_vals_mat_ho[,1],
  ff_pred = fBands[[1]][,1],
  ff_lower = fBands[[1]][,3],
  ff_upper = fBands[[1]][,2]
)

# Plot of the band and true line
ggplot() +
  geom_vline(xintercept = c(1/3, 2/3),
             color = "lightgrey") +
  geom_ribbon(aes(x = t, ymin = ff_lower, ymax = ff_upper), 
              fill = "#56B4E9", alpha = 0.5, 
              linewidth = 1.25, 
              data = band_plot_df) +
  geom_line(aes(x = t, y = y_true), 
            data = band_plot_df) +
  geom_line(aes(x = t, y = ff_pred), 
             color = "#56B4E9", 
             data = band_plot_df) +
  theme_bw(base_size = 20) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  labs(y = "Functional Response Y(t)", 
       x = "Sampling Point (t)")
  

# Second prediction: c(1,1)
fBands2 <- FunctionalPrediction::predict_concurrent(
  concurrent_list = fReg_list,
  interval = "prediction", 
  err_str = "t",
  new_dat = c(1), 
  conf.level = 0.90, 
  n_int = 3,
  nu0_hat = "singh_df",
  mse_scalar = "ub"
)

# Data frame to plot the band
band_plot_df2 <- data.frame(
  t = seq(0,1, length.out = 101),
  y_true = y_vals_mat_ho[,2],
  ff_pred = fBands2[[1]][,1],
  ff_lower = fBands2[[1]][,3],
  ff_upper = fBands2[[1]][,2]
)

# Plot of the band and true line
ggplot() +
  geom_vline(xintercept = c(1/3, 2/3),
             color = "lightgrey") +
  geom_ribbon(aes(x = t, ymin = ff_lower, ymax = ff_upper), 
              fill = "#56B4E9", alpha = 0.5, 
              linewidth = 1.25, 
              data = band_plot_df2) +
  geom_line(aes(x = t, y = y_true), 
            data = band_plot_df2) +
  geom_line(aes(x = t, y = ff_pred), 
             color = "#56B4E9", 
             data = band_plot_df2) +
  theme_bw(base_size = 20) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) + 
  labs(y = "Functional Response Y(t)", 
       x = "Sampling Point (t)")

```